Gravitational waves from colliding black holes

### 2017-05-03 Report/Draft/draft.tex
Report latest

\documentclass[11pt]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\bibliographystyle{unsrt}
\textwidth=430pt
\textheight=640pt
\hoffset=-50pt
\voffset=-80pt
%opening
\title{Gravitational waves from colliding black holes}
\author{Chris Pedersen}

@@

The qualitative idea behind current methods of parameter estimation is that the GW signal is processed and extracted from the raw strain data, and then matched against an array of simulated waveforms to find which waveform most closely resembles the detected signal. The two technical challenges here are quantifying how well a template represents the observed signal, and how to efficiently sample the parameter space for new templates to test against the data. Within LIGO a standard method for quantifying the similarity between two signals $h_1(t)$ and $h_2(t)$ is given by the noise weighted inner product, also referred to as the 'match' between signals, and is defined as
\begin{equation}
\langle h_1(t) \vert h_2(t) \rangle=2\int_{0}^{\infty}\frac{\tilde{h_1^*}(f)\tilde{h_2}(f)+\tilde{h_1}(f)\tilde{h_2^*}(f)}{S_n(f)}df 
\end{equation}
where $\tilde{h}(f)$ is the Fourier transform of the signal $h(t)$, and $S_n(f)$ is the noise curve of the detector\cite{oldest_pe}. It is also convenient here to define the term \textit{mismatch}, which is simply $1-m$ where $m$ is the match. A variety of methods have been employed to sample the parameter space, including nested sampling\cite{pe2} and analysis using Gaussian wavelets\cite{props}, and in this work we use a framework of Bayesian inference and Markov-Chain Monte Carlo methods\cite{inj}\cite{pe}\cite{pe4}. This process results in a set of posterior distribution functions (PDFs) for each parameter. Using Bayes' Theorem, the posterior is given by
\begin{equation}
p(\vec{\theta})|d)=\frac{p(\vec{\theta})p(d|\vec{\theta})}{P(d)}
\end{equation}
where $\vec{\theta}$ is an $n$ dimensional vector in our parameter space of $n$ parameters\cite{pe2}. The first term in the numerator, $p(\vec{\theta})$ is known as the \textit{prior} distribution, and quantifies knowledge we already have about the system that should influence our estimation of its' parameters. In practice, in inference of BBHs, the priors are almost always uniform or isotropic distributions. The denominator, $P(d)$, is essentially a normalisation factor ensuring that the posterior distribution integrats to unity. The crucial term is the \textit{likelihood} for a given set of parameters given the data, which is determined by
\begin{equation}
p(d|\vec{\theta})\propto \exp\bigg(-\frac{1}{2}\sum_{k=1,2}\Big \langle h^M_k(\vec{\theta})-d_k \big\vert h^M_k(\vec{\theta})-d_k \Big \rangle\bigg)
\end{equation}
where the inner product is that defined in (13), $d_k$ is the observed strain in the detector, and $h_k^M(t;\vec{\theta})$ is the simulated detector response for a given set of parameters $\vec{\theta}$ given by (9).
Various methods of waveform generation (known as \textit{approximants}) have been attempted, with varying degrees of computational intensity. In general, Post-Newtonian expansions are used in the inspiral phase where the gravitational field is weak enough for approximations to be sufficient. During the merger and ringdown, full numerical relativity simulations are required as the curvature is sufficiently strong that Post-Newtonian approximations are no longer valid\cite{waveforms}\cite{imr}. These require significantly more computation time, and as such it is only in recent years that waveforms describing the full inspiral, merger and ringdown phases have become available.

--- some equation for SNR
--- overview of MCMCs that will be used for inference

@@

In this section we intend to get some idea of the areas in which we expect the detector to be most sensitive to precession, and those in which precessing signals are intrinsically degenerate with non-precessing waveforms. This is done by finding the match between waveforms with and without in-plane spin for a range of parameters. In the case of a strong match ($>0.95$), it is unlikely that $chi_p$ can be accurately and precisely recovered for signals with those parameters due to an intrinsic similarity of the waveforms.

\begin{figure}
	\includegraphics[scale=0.55]{fig10.png}
	\centering
	\caption{Matches between precessing and non-precessing waveforms for a range of inclinations and mass ratios.(make this more compact before submit)}
	\centering
\end{figure}
\begin{figure}
	\includegraphics[scale=0.55]{fig11.png}
	\centering
	\caption{Spin difference required for the match between precessing and non-precessing waveforms to drop below 0.95}
	\centering
\end{figure}

Fig. 10 shows how the matches between precessing and non-precessing waveforms change with inclination for four different mass ratios. The dark regions of the match is where the signals have significant mismatch, and the waveforms should be well distinguished, and hte light regions are where the waveforms are effectively degenerate. It is apparent that the degeneracy is far stronger for close to equal mass binaries, where only maximally precessing systems at edge-on inclinations will have noticeably different waveforms. Here we plot the 0.97 match contour as it is generally expected that the advanced LIGO detectors should be able to distinguish between waveforms with a match of 0.97. For all but the most extreme mass ratios, a combination of significant in-plane spin ($/geq0.6$) and/or strong inclination is necessary to have any significant deviation for systems with high precession. It is important to remember here that as we change the inclination, the overall strength of the signal will drop (with strongest at $\iota=0$), and the matching process does not account for this change in signal strength.

@@

\begin{figure}
	\includegraphics[scale=0.85]{fig12.png}
	\centering
	\caption{Match limit plots for a range of inclinations and phases for 4 polarisations.}
	\centering
\end{figure}

@@

\section{Detector response simulations}
In this section we describe the process of inserting signals into simulated detector noise, and then perform a full parameter estimation.
\subsection{Signal extraction}
Data whitening - basically LOSC stuff, how we go from raw data to a signal - move this to intro?
\subsection{Inference pipeline}
A large number of software injections were going to be necessary to generate any interesting results due to the large size and complexity of the parameter space as revealed by the match results. So it was key that they could be performed efficiently and in an organised way. This required the construction of an inference pipeline which formed the bulk of the computational work of the project. This was ultimately one bash script that would create a new folder for each run, generate the injection file and run the inference MCMC.  It also saves the injected parameters as a python dictionary, plots all posteriors overlaid with injected parameters, plots the injected waveform, and the recovered (maximum posterior, or \textit{MAP} waveform) and injected waveforms overlaid on the whitened detector strain data, and return matches between the MAP and injected waveforms. This meant that all the relevant analysis and data processing needed for each inference run was fully automated. A flow chart of this process is shown in Fig. 12. For the MCMC parameters, in order to acheive a balance between computation time and result accuracy, 5,000 samplers were used and 12,000 iterations, and the burn-in process was skipped. These parameters were kept the same for all software injections.
This project work was also split between a number of different LIGO clusters, as well as work on several different local PCs, and so the pipeline along with all the relevant match and precession scripts were maintained as part of a GitHub repository which meant that the codes could be developed effectively and without conflict.
\begin{figure}
	\includegraphics[scale=0.55]{fig13.png}
	\centering
	\caption{Flow chart of the inference pipeline. The ellipses represent data files that were generated during the pipeline and stored in case further analysis was necessary. The boxes at the bottom represent the final output of the pipeline.}
	\centering
\end{figure}
\subsection{Inference results}
In this section we present the results from a range of inference runs, attempt to assess the effectiveness of spin estimation and compare the inference results with the match results to evaluate how effective the match process is as a predictor of inference accuracy. For all injections we set $S_{1\perp}=0.9$ and $|\vec{S_2}|=0$ with $\chi_p=0.9$, and cover a uniform range of inclinations. We set the phase to be $\phi=0.$, and the polarisation angle $\psi=0.8$, so the signals are of mixed polarisation, but slightly cross dominated. The match results indicate that cross polarised signals generally have a higher spin resolution, but we want to avoid the signal drop at $\iota=\pi/2$, which is why a mixed but cross dominated polarisation angle was selected. The source location and coalesence time were also kept fixed for all injections. We also only examine high mass ratio binaries, as the match results indicate that precession is not significant in equal mass ratio binaries. Out of a total of ****** software injections that were performed, only **** instances where the injected $\chi_p$ value fell within the 90\% confidence interval. In Figs 14 and 15 we compare posteriors for two inferences, where $\chi_p$ was excluded and where it was recovered moderately well, and we refer to these as A and B respectively. Even in the cases where the injected value was within the 90\% confidence intervals, it was always closer to the upper bound than to the mean, and a mean $\chi_p$ greater than 0.65 was never found.

\begin{figure}
	\includegraphics[scale=1]{fig14.png}
	\centering
	\caption{Posteriors for an injection with $m1=50$, $m2=15$, $\iota=2.51$, $\phi=0$, $\psi=0.8$ at a distance of 200Mpc. Prior distributions are included for the derived spin parameters $\chi_p$ and $\chi_{eff}$, which result from uniform spin magnitude and isotropic orientation distributions.}
	\centering
\end{figure}
\begin{figure}
	\includegraphics[scale=1]{fig15.png}
	\centering
	\caption{Posteriors for an injection with the same parameters as in Fig. 14, expect here with an inclination of $\iota=0.63$}
	\centering
\end{figure}

@@

\begin{figure}[h]
	\includegraphics[scale=0.3]{fig16.png}
	\centering
	\caption{Strain plots for both H1 and L1 detectors. The top two graphs show the strain data corresponding to the results in Fig. 14, the bottom two graphs correspond to Fig. 15. The green and red lines show the whitened strain data, the black line is the MAP waveform and the injected waveform is shown in blue.}
	\centering
\end{figure}

The results shown in Fig. 16 show a stronger match for A, where the majority of parameters were more accurately recovered. In particular, for B, L1 shows a low match of 0.93 between MAP and injected waveforms. This is likely due to the precessive effects which are most apparent around $~-0.4s$, where the injected signal is almost entirely flat, and later on during the inspiral around $~-0.2s$ where there is noteable disagreement between the signals. In this case, it is clear that the inference process has not found an optimal template waveform to replicate the injected signal. However in the case of A there is a strong match between waveforms, and it is clear from the strain plot that the MAP waveform matches the strain data extremely well. There is, however, a mild disagreement during the late merger and ringdown, which is the source of the mismatches of 0.01 and 0.02. Since the match is so strong, the waveforms themselves must be intrinsically degenerate up to deviation that occurs at the merger.

@@

From Fig. 18 it can be seen that the amplitude of the signal noticeably drops towards the end of the inspiral and just before the merger, implying that this would be an ideal signal to recover a high value of $\chi_p$ from. The posteriors shown in Fig. 17 are mostly accurate, with the majority of injected parameters lying inside the 90\% confidence intervals, however once again $\chi_p$ is excluded, and significantly underestimated. The match values shown in Fig. 18 are strong, with both detectors giving a match of 0.96 between the injected and MAP waveforms. This is further evidence of an intrinsic degeneracy that prevents high $\chi_p$ MAP values from being found. For all posteriors shown, both $\phi$ and $\psi$ are broadly uniform, and in the case where they are not uniform, they are largely inaccurate. However in section 2 it was observed that $\psi$ and especially $\phi$ have a significant effect on the structure of the parameter space, and the match between precessing and non-precessing waveforms. One possible explanation for this uniformity is that for each stochastically selected phase and polarisation, the MCMC algorithm readily finds a good match between the injected signal and the strain data simply by tweaking the other remaining parameters slightly. If this is the case, since there are a larger number of waveforms in the template bank with lower values of $\chi_p$ (as shown in the prior distributions), if these waveforms can still match the signal well just by slightly changing inclination, source location and masses, we would expect the MCMC algorithm to favour lower values of $\chi_p$ but with slightly larger uncertanties on the other posteriors. This explanation is exemplified by the waveforms in Fig. 8, where it is shown that the magnitude of the precessive effect is not necessarily proportional to $\chi_p$. It is also interesting to note that $\chi_{eff}$ parameters are generally relatively accurate, which implies it is mainly spin magnitude that is being underestimated, but the orientations are recovered accurately.
\begin{figure}
	\includegraphics[scale=1]{fig17.png}
	\centering
	\caption{Posteriors for injection C, which has the same parameters as in Fig. 14, but with an inclination of $\iota=1.26$ and at a distance of 800Mpc.}
	\centering
\end{figure}
\begin{figure}
	\includegraphics[scale=0.9]{fig18.png}
	\centering
	\caption{Strain plot of injection C, a waveform with strong precessive modulation.}
	\centering
\end{figure}
\subsection{Inclination and precession}
Next we examine the effect of inclination on $\chi_p$ recovery. Software injections were performed for a range of inclinations at two distances, 200Mpc and 800Mpc, which simulate a strong and weak signal respectively.
\subsection{Impact of Virgo}
A look at how Virgo will influence PE, especially chi p estimation.
\section{Conclusions}
Summary of results on PE and estimation of chi p, and a prospect on Virgo's impact.
\section{Reflective Statement}
I now need a therapist

@@

1. SNR equation
2. MCMC description
3. Match plots for phase and inclination
4. Inference results - 2D posteriors
4. Inference results - strain plots
4. Inference results - Violin plots
5. Virgo results
6. Conclusions
7. Proper references of pics

>>>>

\documentclass[11pt]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}
\bibliographystyle{unsrt}
\textwidth=430pt
\textheight=640pt
\hoffset=-50pt
\voffset=-80pt
%opening
\title{Gravitational waves from colliding black holes}
\author{Chris Pedersen}

@@

The qualitative idea behind current methods of parameter estimation is that the GW signal is processed and extracted from the raw strain data, and then matched against an array of simulated waveforms to find which waveform most closely resembles the detected signal. The two technical challenges here are quantifying how well a template represents the observed signal, and how to efficiently sample the parameter space for new templates to test against the data. Within LIGO a standard method for quantifying the similarity between two signals $h_1(t)$ and $h_2(t)$ is given by the noise weighted inner product, also referred to as the 'match' between signals, and is defined as
\begin{equation}
\langle h_1(t) \vert h_2(t) \rangle=2\int_{0}^{\infty}\frac{\tilde{h_1^*}(f)\tilde{h_2}(f)+\tilde{h_1}(f)\tilde{h_2^*}(f)}{S_n(f)}df 
\end{equation}
where $\tilde{h}(f)$ is the Fourier transform of the signal $h(t)$, and $S_n(f)$ is the noise curve of the detector\cite{oldest_pe}. It is also convenient here to define the term \textit{mismatch}, which is simply $1-m$ where $m$ is the match. A variety of methods have been employed to sample the parameter space, including nested sampling\cite{pe2} and analysis using Gaussian wavelets\cite{props}, and in this work we use a framework of Bayesian inference and Markov-Chain Monte Carlo methods\cite{inj}\cite{pe}\cite{pe4}. This process results in a set of posterior distribution functions (PDFs) for each parameter. Using Bayes' Theorem, the posterior is given by
\begin{equation}
p(\vec{\theta})|d)=\frac{p(\vec{\theta})p(d|\vec{\theta})}{P(d)}
\end{equation}
where $\vec{\theta}$ is an $n$ dimensional vector in our parameter space of $n$ parameters\cite{pe2}. The first term in the numerator, $p(\vec{\theta})$ is known as the \textit{prior} distribution, and quantifies knowledge we already have about the system that should influence our estimation of its' parameters. In practice, in inference of BBHs, the priors are almost always uniform or isotropic distributions. The denominator, $P(d)$, is essentially a normalisation factor ensuring that the posterior distribution integrats to unity. The crucial term is the \textit{likelihood} for a given set of parameters given the data, which is determined by
\begin{equation}
p(d|\vec{\theta})\propto \exp\bigg(-\frac{1}{2}\sum_{k=1,2}\Big \langle h^M_k(\vec{\theta})-d_k \big\vert h^M_k(\vec{\theta})-d_k \Big \rangle\bigg)
\end{equation}
where the inner product is that defined in (13), $d_k$ is the observed strain in the detector, and $h_k^M(t;\vec{\theta})$ is the simulated detector response for a given set of parameters $\vec{\theta}$ given by (9). The SNR is given by

\begin{equation}
\mathrm{SNR}=\sqrt{\sum_{\mathrm{det}}^{}\int_{f_{\mathrm{low}}}^{f_{\mathrm{high}}}\frac{|h_{\mathrm{det}}(f,\vec{\theta})|}{S_{\mathrm{det}}(f)}\mathrm{d}f}
\end{equation}

\noindent where the sum is over the detectors, and $h_{\mathrm{det}}(f,\vec{\theta})$ is the signal in the detector, and $S_{\mathrm{det}}(f)$ is the noise. The posterior distributions are sampled stochastically using a Markov-Chain Monte Carlo process and the Metropolis-Hasting algorithm. The qualitative idea behind this is that a sampler random walks across the parameter space. At each iteration, a random step in the parameter space is proposed, and the posterior is sampled at this new point. If the posterior has a higher value, this new point is used as the starting point for the next iteration. However if the posterior is lower, then the sampler only has a probability to jump to this new step proportional to the ratio of the posteriors at the current and proposed steps. In this way, after each iteration the distribution of samplers gets a step closer to accurately representing the posterior.

Various methods of waveform generation (known as \textit{approximants}) have been employed, with varying degrees of computational intensity. In general, Post-Newtonian expansions are used in the inspiral phase where the gravitational field is weak enough for approximations to be sufficient. During the merger and ringdown, full numerical relativity simulations are required as the curvature is sufficiently strong that Post-Newtonian approximations are no longer valid\cite{waveforms}\cite{imr}. These require significantly more computation time, and as such it is only in recent years that waveforms describing the full inspiral, merger and ringdown phases have become available.

@@

In this section we intend to get some idea of the areas in which we expect the detector to be most sensitive to precession, and those in which precessing signals are intrinsically degenerate with non-precessing waveforms. This is done by finding the match between waveforms with and without in-plane spin for a range of parameters. In the case of a strong match ($>0.95$), it is unlikely that $\chi_p$ can be accurately and precisely recovered for signals with those parameters due to an intrinsic similarity of the waveforms.

\begin{figure}
	\includegraphics[width=1\textwidth]{fig10.png}
	\centering
	\caption{Matches between precessing and non-precessing waveforms for a range of inclinations and mass ratios.(make this more compact before submit)}
	\centering
\end{figure}
\begin{figure}
	\includegraphics[width=1\textwidth]{fig11.png}
	\centering
	\caption{Spin difference required for the match between precessing and non-precessing waveforms to drop below 0.95}
	\centering
\end{figure}

Fig. 10 shows how the matches between precessing and non-precessing waveforms change with inclination for four different mass ratios. The dark regions of the match is where the signals have significant mismatch, and the waveforms should be well distinguished, and hte light regions are where the waveforms are effectively degenerate. It is apparent that the degeneracy is far stronger for close to equal mass binaries, where only maximally precessing systems at edge-on inclinations will have noticeably different waveforms. Here we plot the 0.97 match contour as it is generally expected that the advanced LIGO detectors should be able to distinguish between waveforms with a match of 0.97. For all but the most extreme mass ratios, a combination of significant in-plane spin ($\geq0.6$) and/or strong inclination is necessary to have any significant deviation for systems with high precession. It is important to remember here that as we change the inclination, the overall strength of the signal will drop as the amplitude of the gravitational wave is lower (with the peak at $\iota=0$), and the matching process does not account for this change in signal strength.

@@

\begin{figure}
	\includegraphics[width=1\textwidth]{fig12.png}
	\centering
	\caption{Match limit plots for a range of inclinations and phases for 4 polarisations.}
	\centering
\end{figure}

@@

\section{Detector response simulations}
In this section we describe the process of inserting signals into simulated detector noise, and then perform a full parameter estimation.
\subsection{Signal extraction}
Data whitening - basically LOSC stuff, how we go from raw data to a signal - move this to intro?
\subsection{Inference pipeline}
A large number of software injections were going to be necessary to generate any interesting results due to the large size and complexity of the parameter space as revealed by the match results. So it was key that they could be performed efficiently and in an organised way. This required the construction of an inference pipeline which formed the bulk of the computational work of the project. This was ultimately one bash script that would create a new folder for each run, generate the injection file and run the inference MCMC.  It also saves the injected parameters as a python dictionary, plots all posteriors overlaid with injected parameters, plots the injected waveform, and the recovered (maximum posterior, or \textit{MAP} waveform) and injected waveforms overlaid on the whitened detector strain data, and return matches between the MAP and injected waveforms. This meant that all the relevant analysis and data processing needed for each inference run was fully automated. A flow chart of this process is shown in Fig. 12. For the MCMC parameters, in order to acheive a balance between computation time and result accuracy, 5,000 samplers were used and 12,000 iterations, and the burn-in process was skipped. These parameters were kept the same for all software injections.
This project work was also split between a number of different LIGO clusters, as well as work on several different local PCs, and so the pipeline along with all the relevant match and precession scripts were maintained as part of a GitHub repository which meant that the codes could be developed effectively and without conflict.
	\includegraphics[scale=0.55]{fig13.png}
	\caption{Flow chart of the inference pipeline. The ellipses represent data files that were generated during the pipeline and stored in case further analysis was necessary. The boxes at the bottom represent the final output of the pipeline.}
\subsection{Inference results}
In this section we present the results from a range of inference runs, attempt to assess the effectiveness of spin estimation and compare the inference results with the match results to evaluate how effective the match process is as a predictor of inference accuracy. For all injections we set $S_{1\perp}=0.9$ and $|\vec{S_2}|=0$ with $\chi_p=0.9$, and cover a uniform range of inclinations. We set the phase to be $\phi=0.$, and the polarisation angle $\psi=0.8$, so the signals are of mixed polarisation, but slightly cross dominated. The match results indicate that cross polarised signals generally have a higher spin resolution, but we want to avoid the signal drop at $\iota=\pi/2$, which is why a mixed but cross dominated polarisation angle was selected. The source location and coalesence time were also kept fixed for all injections. We also only examine high mass ratio binaries, as the match results indicate that precession is not significant in equal mass ratio binaries. Prior distributions are all uniform, with mass ranging $[5,80]$ and distance in the range $[100,1000]$, and sky location and spin orientations all have isotropic priors. Out of a total of ****** software injections that were performed, there were only **** instances where the injected $\chi_p$ value fell within the 90\% confidence interval. In Figs 14 and 15 we compare posteriors for two inferences, where $\chi_p$ was excluded and where it was recovered moderately well, and we refer to these as A and B respectively. Even in the cases where the injected value was within the 90\% confidence intervals, it was always closer to the upper bound than to the mean, and a mean $\chi_p$ greater than 0.65 was never found.

\begin{figure}
	\includegraphics[width=1\textwidth]{fig14.png}
	\centering
	\caption{Posteriors for an injection with $m1=50$, $m2=15$, $\iota=2.51$, $\phi=0$, $\psi=0.8$ at a distance of 200Mpc. Prior distributions are included for the derived spin parameters $\chi_p$ and $\chi_{eff}$, which result from uniform spin magnitude and isotropic orientation distributions. The red bar shows the injected value, with the black solid line showing the mean of the posterior, and the black dashed line showing the 90\% confidence intervals, which are the standard error margins quoted on LIGO inference results.}
	\centering
\end{figure}
\begin{figure}
	\includegraphics[width=1\textwidth]{fig15.png}
	\centering
	\caption{Posteriors for an injection with the same parameters as in Fig. 14, except here with an inclination of $\iota=0.63$}
	\centering
\end{figure}
\begin{figure}[h]
	\includegraphics[scale=0.3]{fig16.png}
	\centering
	\caption{Strain plots for both H1 and L1 detectors. The top two graphs show the strain data corresponding to the results in Fig. 14, the bottom two graphs correspond to Fig. 15. The green and red lines show the whitened strain data, the black line is the MAP waveform and the injected waveform is shown in blue.}
	\centering
\end{figure}

@@

The results shown in Fig. 16 show a stronger match for A, where the majority of parameters were more accurately recovered. In particular, for B, L1 shows a low match of 0.93 between MAP and injected waveforms. This is likely due to the precessive effects which are most apparent around $-0.4s$, where the injected signal is almost entirely flat, and later on during the inspiral around $-0.2s$ where there is noteable disagreement between the signals. In this case, it is clear that the inference process has not found an optimal template waveform to replicate the injected signal. However in the case of A there is a strong match between waveforms, and it is clear from the strain plot that the MAP waveform matches the strain data extremely well. There is, however, a mild disagreement during the late merger and ringdown, which is the source of the mismatches of 0.01 and 0.02. Since the match is so strong, the waveforms themselves must be intrinsically degenerate up to deviation that occurs at the merger.

@@

\begin{figure}[h]
	\includegraphics[width=1\textwidth]{fig17.png}
	\centering
	\caption{Posteriors for injection C, which has the same parameters as in Fig. 14, but with an inclination of $\iota=1.26$ and at a distance of 800Mpc.}
	\centering
\end{figure}
\begin{figure}[h]
	\includegraphics[scale=0.3]{fig18.png}
	\centering
	\caption{Strain plot of injection C, a waveform with strong precessive modulation.}
	\centering
\end{figure}

From Fig. 18 it can be seen that the amplitude of the signal noticeably drops towards the end of the inspiral and just before the merger, implying that this would be an ideal signal to recover a high value of $\chi_p$ from. The posteriors shown in Fig. 17 are mostly accurate, with the majority of injected parameters lying inside the 90\% confidence intervals, however once again $\chi_p$ is excluded, and significantly underestimated. The match values shown in Fig. 18 are strong, with both detectors giving a match of 0.96 between the injected and MAP waveforms. This is further evidence of an intrinsic degeneracy that prevents high $\chi_p$ MAP values from being found. For all posteriors shown, both $\phi$ and $\psi$ are broadly uniform, and in the case where they are not uniform, they are largely inaccurate. However in section 2 it was observed that $\psi$ and especially $\phi$ have a significant effect on the structure of the parameter space, and the match between precessing and non-precessing waveforms.

One possible explanation for this uniformity is that for each stochastically selected phase and polarisation, the MCMC algorithm readily finds a good match between the injected signal and the strain data simply by tweaking the other remaining parameters slightly. If this is the case, since there are a larger number of waveforms in the template bank with lower values of $\chi_p$ (as shown in the prior distributions), if these waveforms can still match the signal well just by slightly changing inclination, source location and masses, we would expect the MCMC algorithm to favour lower values of $\chi_p$ but with slightly larger uncertanties on the other posteriors. This explanation is exemplified by the waveforms in Fig. 8, where it is shown that the magnitude of the precessive effect is not necessarily proportional to $\chi_p$. It is also interesting to note that $\chi_{eff}$ parameters are generally relatively accurate, which implies it is mainly spin magnitude that is being underestimated, but the orientations are recovered accurately.

\subsection{Inclination and precession}

Next we examine the effect of inclination on $\chi_p$ recovery. Software injections were performed for a range of inclinations at two distances, 200Mpc and 800Mpc, which simulate a strong and weak signal respectively. Posteriors for $\chi_p$ and $\chi_{eff}$ are shown in Fig. 19 for 6 inclinations in the range $[0,\pi]$, with the injected values shown by the red bar. Interestingly there is no apparent correlation between the trends for $\chi_p$ at 200Mpc and at 800Mpc. If anything the trends appear to be opposite, with the most greater accuracy at face-on for the signals at 200Mpc, and greatest accuracy close to edge-on for the 800Mpc signals. As in the previous results, $\chi_{eff}$ is recovered consistently well with the exception of two results around $\iota\approx2$ in the 200Mpc signals, again implying that spin orientation is accurate, but magnitude is underestimated. Therefore there appears to be no clear correlation between $\chi_p$ and $\chi_{eff}$ accuracy.
\begin{figure}[h]
	\includegraphics[width=1\textwidth]{fig19.png}
	\centering
	\caption{Comparisons of derived spin parameters as a function inclination for strong and weak signals. All other parameters are kept the same as listed in Fig. 14.}
	\centering
\end{figure}
\begin{figure}[h]
	\includegraphics[scale=0.6]{fig20.png}
	\centering
	\caption{Match between MAP and injected waveforms along with spin inaccuracy for the 200Mpc and 800Mpc runs as a function of inclination.}
	\centering
\end{figure}

In Fig. 20 we attempt to identify how well the posteriors in Fig. 19 correlate with the predicted trends from the match results in Figs. 10 and 11. If the match process were a reliable method of predicting inference results, the posteriors for $\chi_p$ around inclinations of 1 to 2 should be the most accurate, as precessing waveforms are most distinct from precessing ones here. However the 200Mpc results in fact show the opposite trend, and accuracy drops to a minimum around this area. 
\subsection{Impact of Virgo}
A look at how Virgo will influence PE, especially chi p estimation.
\begin{figure}[h]
	\includegraphics[width=1\textwidth]{fig21.png}
	\centering
	\caption{Plots showing $\chi_p$ and $\chi_{eff}$ for a range of inclinations at 200Mpc with the VIRGO detector included in the inference process.}
	\centering
\end{figure}
\section{Conclusions}
Summary of results on PE and estimation of chi p, and a prospect on Virgo's impact.
\section{Reflective Statement}
I now need a therapist

@@

6. Conclusions
7. Proper references of pics
